---
title: "VaxGO"
author: "Wasim Syed (@wasimvacinas)"
output: 
  flexdashboard::flex_dashboard:
    logo: "vaxgo_logo_40.png"
    orientation: columns
    theme:
      bg: "#ffffff"
      fg: "#000000" 
      primary: "#ffffff"
      base_font:
        google: Montserrat
      code_font:
        google: Montserrat
    source_code: embed 
    navbar:
        - { icon: "fa-instagram", href: "https://instagram.com/wasimvacinas", align: right}
        - { icon: "fa-linkedin", href: "https://www.linkedin.com/in/wapsyed", align: right}
        - { icon: "ion-github", href: "github.com/wapsyed", align: right}
runtime: shiny
---

<style>
  body {
    font-size: 12px;
  }
  
  .form-group,
  .shiny-input-container {
    font-size: 12px; 
  }
  
  .shiny-input-container input,
  .shiny-input-container select,
  .shiny-input-container textarea {
    font-size: 12px;
  }
  
  .btn,
  .btn-primary {
    font-size: 12px; 
  }
  
  .table,
  .table th,
  .table td {
    font-size: 12px; 
  }
  
  /* Ajuste para as abas */
  .nav-tabs > li > a {
    font-size: 12px !important; 
  }

  .nav-tabs > li.active > a {
    font-size: 12px !important; 
  }

  .nav-tabs > li > a:hover {
    font-size: 12px !important; 
  }

  .tab-content {
    font-size: 12px !important;
  }
</style>

```{r setup, include=FALSE}
# Install packages ------
# install.packages("flexdashboard")
# install.packages("datasets")
# install.packages("knitr")
# devtools::install_github("rstudio/d3heatmap")
# install_github("jokergoo/InteractiveComplexHeatmap")
# install.packages("biclust")
# install.packages("BH")
# install.packages('heatmaply')
# install.packages("devtools")
# Bioconductor packages ------------------------
# install.packages("BiocManager")
# if (!require("BiocManager", quietly = TRUE))
#   install.packages("BiocManager")

# BiocManager::install("GEOquery")
# BiocManager::install("DESeq2")
# BiocManager::install("biomaRt")
# BiocManager::install("celldex")
# BiocManager::install("GSVA")
# BiocManager::install("sva")
# BiocManager::install("clusterProfiler")
# BiocManager::install("msigdbr")
# BiocManager::install("EnhancedVolcano")
# BiocManager::install("AnnotationDbi")
# BiocManager::install("ComplexHeatmap")
# BiocManager::install("InteractiveComplexHeatmap")
# BiocManager::install("cytolib")
# BiocManager::install("ImmuneSpaceR")
# BiocManager::install("org.Hs.eg.db", character.only = TRUE)

# Call library ------
library(BiocManager)
# library(devtools)
# library(GEOquery)
# library(Matrix)
# library(circlize)
# library(RColorBrewer)
# library(celldex)
# library(biomaRt)
# library(EnhancedVolcano)
# library(ggbeeswarm)
# library(corrr)
# library(ggcorrplot)
# library(FactoMineR)
# library(factoextra)
# library(esquisse)
# library(ComplexHeatmap)
# library(msigdbr)
# library(ape)
# library(GSVA)
# library(sva)
# library(patchwork)
# library(extrafont)
# library(showtext)
# library(ggheatmap)
# library(InteractiveComplexHeatmap)
# library(ggstatsplot)
# library(datasets)
# library(InteractiveComplexHeatmap)
# library(biclust)
# library(ImmuneSpaceR)
library(org.Hs.eg.db)
library(plotly)
library(DESeq2)
library(clusterProfiler)
library(janitor)
library(ggsci)
library(ggthemes)
library(here)
library(readxl)
library(scales)
library(flexdashboard)
library(vegan)
library(corto)
library(plotly)
library(flexdashboard)
library(DT)
library(org.Hs.eg.db)
library(BH)
library(tidyverse)
library(cytolib)
library(heatmaply)



#Increase upload file size
options(shiny.maxRequestSize=30*1024^2)

################### Custom Functions -------
##### autoORA (for multiple conditions)
autoORA <- function(df,
                    TERM2GENE,
                    pAdjustMethod, 
                    pvalueCutoff_ora, 
                    qvalueCutoff_ora,
                    geneset_name) {
  
  enricher_conditional <- function(condition_2, df, TERM2GENE, pAdjustMethod, pvalueCutoff_ora, qvalueCutoff_ora) {
    genes <- df %>% 
      filter(condition == condition_2) %>% 
      pull(genes)
    
    tryCatch({
      enricher(gene = genes,
               TERM2GENE = TERM2GENE,
               pAdjustMethod = pAdjustMethod,
               pvalueCutoff = pvalueCutoff_ora, 
               qvalueCutoff = qvalueCutoff_ora) %>% 
        as.data.frame() %>%
        mutate(condition = condition_2,
               geneset_name = geneset_name)
    }, error = function(e) {
      NULL
    })
  }
  
  ora_results <- df %>%
    distinct(condition) %>%
    pull(condition) %>%
    map_dfr(~ enricher_conditional(.x, df, TERM2GENE, pAdjustMethod, pvalueCutoff_ora, qvalueCutoff_ora))
  
  ora_results_final <- ora_results %>%
    rownames_to_column("delete") %>%
    dplyr::select(-delete, -Description) %>%
    dplyr::select(condition, geneset_name, everything()) %>% 
    clean_names() %>% 
    rename(process = id)
  
  return(ora_results_final)
}


####### autoGSEA (for multiple conditions)
autoGSEA <- function(df, TERM2GENE, 
                     geneset_name, 
                     minGSSize, 
                     maxGSSize,
                     pvalueCutoff,
                     pAdjustMethod) {
  results <- list()
  
  condicoes <- df %>%
    dplyr::pull(condition) %>%
    unique() %>%
    as.character()
  
  for (condicao in condicoes) {
    degs_condicao <- df %>%
      dplyr::filter(condition == condicao) %>%
      dplyr::select(genes, log2fold_change) %>%
      distinct() %>%
      mutate(rank = rank(log2fold_change, ties.method = "random")) %>%
      arrange(desc(rank))
    
    gene_list_lf2c <- as.vector(degs_condicao$log2fold_change)
    names(gene_list_lf2c) <- degs_condicao$genes
    gene_list_lf2c <- na.omit(gene_list_lf2c)
    gene_list_lf2c <- sort(gene_list_lf2c, decreasing = TRUE)
    
    auto_gsea <- tryCatch({
      GSEA(geneList = gene_list_lf2c,
           TERM2GENE = TERM2GENE,
           minGSSize = minGSSize,
           maxGSSize = maxGSSize,
           pvalueCutoff = pvalueCutoff,
           pAdjustMethod = pAdjustMethod) %>%
        as.data.frame() %>%
        arrange(qvalue) %>%
        mutate(condition = condicao,
               gsea_enrichment = geneset_name)
    }, error = function(e) {
      return(NULL)
    })
    
    if (!is.null(auto_gsea)) {
      results[[paste(condicao, geneset_name, sep = "_")]] <- auto_gsea
    }
  }
  
  final_result <- bind_rows(results) %>%
    rename(process = ID) %>%
    dplyr::select(-Description) %>%
    clean_names() %>%
    dplyr::select(condition, everything())
  
  return(final_result)
}


# Function for overlapping -------
overlap_genes <- function(cond1, cond2, data) {
  genes_cond1 <- data$genes[data$process == cond1]
  genes_cond2 <- data$genes[data$process == cond2]
  
  genes_shared <- intersect(genes_cond1, genes_cond2)
  
  genes_notshared_cond1 <- setdiff(genes_cond1, genes_cond2)
  genes_notshared_cond2 <- setdiff(genes_cond2, genes_cond1)
  
  total_genes_cond1 <- length(genes_cond1)
  total_genes_cond2 <- length(genes_cond2)
  
  percentage_shared_cond1 <- length(genes_shared) / total_genes_cond1 * 100
  percentage_shared_cond2 <- length(genes_shared) / total_genes_cond2 * 100
  
  shared_genes <- data.frame(
    Cond1 = cond1,
    Cond2 = cond2,
    Shared = length(genes_shared),
    NotShared_cond1 = length(genes_notshared_cond1),
    NotShared_cond2 = length(genes_notshared_cond2),
    Total_Genes_Cond1 = total_genes_cond1,
    Total_Genes_Cond2 = total_genes_cond2,
    Genes_Names = paste(genes_shared, collapse = ", "),
    Percentage_Shared_Cond1 = percentage_shared_cond1,
    Percentage_Shared_Cond2 = percentage_shared_cond2
  )
  
  return(shared_genes)
}


# Fisher's exact test
fisher_exact_test <- function(shared, notshared1, notshared2) {
  cont_table <- matrix(c(shared, notshared1, notshared2, 0), nrow = 2)
  results_fisher <- fisher.test(cont_table)
  return(results_fisher$p.value)
}
```


```{r global, include=FALSE}
############################### Load data ------

########## Gene sets ------

######## IMMUNEGO -----
# ImmuneGO_Annotated_GO <- read_csv(here("Tables", "ImmuneGO_Annotated_GO_2024-05-28.csv"))
# ImmuneGO_Annotated_GO %>% 
#   saveRDS(here("Tables", "ImmuneGO_Annotated_GO_2024-05-28.rds"))
ImmuneGO_Annotated_GO <- readRDS(here("Tables", "ImmuneGO_Annotated_GO_2024-05-28.rds"))

ImmuneGO_annotation = ImmuneGO_Annotated_GO %>% 
  dplyr::select(process, immune_system:immune_tissue) 

ImmuneGO_genes_general = ImmuneGO_Annotated_GO %>% 
  separate_rows(genes, sep = ",") %>% 
  filter(go_term == "Manual",
         !process %in% c("ADAPTIVE IMMUNE SYSTEM",
                         "INNATE IMMUNE SYSTEM",
                         "BCR REPERTOIRE",
                         "TCR REPERTOIRE")) %>% 
  dplyr::select(process, genes)

ImmuneGO_genes_specific = ImmuneGO_Annotated_GO %>% 
  separate_rows(genes, sep = ",") %>% 
  filter(!go_term == "Manual") %>% 
  dplyr::select(process=gene_set_short, genes)

######## CELL MARKER IMMUNE ---------

# CellMarker_ImmuneCells = read_csv(here("Tables", "CellMarker_ImmuneCells.csv"))
# CellMarker_ImmuneCells %>% 
#   saveRDS(here("Tables", "CellMarker_ImmuneCells.rds"))
CellMarker_ImmuneCells = readRDS(here("Tables", "CellMarker_ImmuneCells.rds"))

CellMarker_annotation = CellMarker_ImmuneCells %>% 
  dplyr::select(cell_name, Type) %>% 
  distinct()

CellMarker_genes = CellMarker_ImmuneCells %>% 
  dplyr::select(process = cell_name, genes = marker)

######## VAX MSIGDB --------
# VaxSigDB_Gene_sets_Annotated_RAW <- read_csv(here("Tables", "VaxSigDB_Gene_sets_Annotated_RAW.csv"))
# VaxSigDB_Gene_sets_Annotated_RAW %>% 
#   saveRDS(here("Tables", "VaxSigDB_Gene_sets_Annotated_RAW.rds"))
VaxSigDB_Gene_sets_Annotated_RAW <- readRDS(here("Tables", "VaxSigDB_Gene_sets_Annotated_RAW.rds"))

#Filters
vaxsig_columns = c("VACCINE", "PLATFORM", "TARGET_PATHOGEN_DISEASE", "MICROBE_TYPE", "STUDY_TYPE", "STUDY_SUBTYPE","DATE-TIME", "DATE", "TIME", "AGE", "AGE_CATEGORY", "SAMPLE_SOURCE", "SYSTEMATIC_NAME", "GENE_SYMBOLS")
filter_vaxsig_samplesource = "PBMC"
filter_vaxsig_studytype = "VACCINE"
filter_vaxsig_studysubtype = "VAC ONLY"

VaxSigDB_Genesets_filtered = VaxSigDB_Gene_sets_Annotated_RAW %>% 
  dplyr::select(vaxsig_columns) %>% 
  filter(SAMPLE_SOURCE == filter_vaxsig_samplesource,
         STUDY_TYPE == filter_vaxsig_studytype,
         STUDY_SUBTYPE == filter_vaxsig_studysubtype) 
#Annotation
VaxSigDB_annotation = VaxSigDB_Genesets_filtered %>% 
  mutate(process = paste0(VACCINE, " (", `DATE-TIME`, ", ", AGE, " YO)")) %>% 
  dplyr::select(process, PLATFORM, TARGET_PATHOGEN_DISEASE, MICROBE_TYPE, DATE, TIME)
#Genes
VaxSigDB_Genes_filtered = VaxSigDB_Genesets_filtered %>% 
  separate_rows(GENE_SYMBOLS, sep = ",") %>% 
  mutate(process = paste0(VACCINE, " (", `DATE-TIME`, ", ", AGE, " YO)")) %>% 
  dplyr::select(process, genes = GENE_SYMBOLS)

######## BTM MODULES --------
# btm_annotation_table <- read_csv(here("Tables", "btm_annotation_immune_table_labelled.csv"))
# btm_annotation_table %>%  saveRDS(here("Tables", "btm_annotation_immune_table_labelled.rds"))
btm_annotation_table <- readRDS(here("Tables", "btm_annotation_immune_table_labelled.rds"))

#Filters
filter_btm_category = "immune"
filter_btm_annotationlevel = c("complete", "partial") #Anyone
# filter_btm_annotationlevel = "complete"
# filter_btm_annotationlevel = "partial"

#Annotation
btm_annotation = btm_annotation_table %>% 
  filter(module_category == filter_btm_category,
         annotation_level %in% filter_btm_annotationlevel) %>% 
  dplyr::select(immune_system, immune_subsystem, cells, composite_name)
#Genes
btm_genes = btm_annotation_table %>% 
  filter(module_category == filter_btm_category,
         annotation_level %in% filter_btm_annotationlevel) %>% 
  dplyr::select(process = composite_name, genes = module_member_genes) %>% 
  separate_rows(genes, sep = ",")

#Vaccine Atlas TERMINAR -----
# study <- CreateConnection("IS2")
# 
# dataset <- study$getDataset("demographics", original_view = FALSE)
# VaxAtlas_all_norm_eset <- readRDS(here("Tables", "VaxAtlas_all_norm_eset.rds"))

```



# DEGs analysis


## Column {.sidebar data-width=300}


**Upload your data**

```{r}
#INPUTS

#https://shiny.posit.co/r/gallery/widgets/widget-gallery/ 

############## Upload files ############## 

textInput(inputId = "filename", 
          label = "Project title", 
          value = "my_condition_degs")


fileInput(inputId = "conditions_genes", 
          label = "DEGs L2FC",
           accept = ".csv")
```

```{r}
# checkboxGroupInput(inputId = "columns_annotation", 
#                    label = "Annotations", 
#                    choices = NULL)
# 
# # Reactive expression to read the uploaded file
# annotations <- reactive({
#   req(input$annotation_conditions)
#   read.csv(input$annotation_conditions$datapath)
# })
# 
# # Observe the file input and update the checkbox group choices
# observe({
#   req(annotations())
#   updateCheckboxGroupInput(session, "columns_annotation",
#                            choices = names(annotations()))
# })
# 
# tableOutput("selected_columns")
```

**Filters**
```{r}
############## Filters 

# DGE analysis
numericInput(inputId = "filter_padj", 
             label = "Adjusted p-value cutoff", 
             value = 0.05, 
             min = 0, max = 1, step = 0.01)

numericInput(inputId = "filter_logfc", 
             label = "Log2 Fold Change cutoff", 
             value = 1, 
             min = 0, max = 10, step = 0.1)

actionButton("go_1", "Go")

```

## Row {.tabset .tabset-fade}

### Volcano plot

```{r}

renderPlotly({
  req(input$go_1) # Action button
  req(conditions_genes())
  
  # Prepare the data
  data <- conditions_genes() %>%
    mutate(log_q = -log10(padj))
  
  # List of unique conditions
  unique_conditions <- unique(data$condition)
  
  # Create an empty plot
  p <- plot_ly()
  
  # Add a trace for each condition
  for (cond in unique_conditions) {
    p <- p %>%
      add_trace(
        data = data[data$condition == cond, ],
        x = ~log2fold_change, 
        y = ~log_q, 
        text = ~genes,
        type = 'scatter', 
        mode = 'markers',
        marker = list(
          color = ~log2fold_change,
          size = 10,
          opacity = 0.9,
          colorscale = c("#3399FF", "white", "#FF6666")
        ),
        name = cond,
        visible = ifelse(cond == unique_conditions[1], TRUE, FALSE) # Show only the first condition by default
      )
  }
  
  # Create updatemenus based on conditions
  condition_buttons <- lapply(1:length(unique_conditions), function(i) {
    visibility <- rep(FALSE, length(unique_conditions))
    visibility[i] <- TRUE
    list(
      method = "restyle",
      args = list("visible", visibility),
      label = unique_conditions[i]
    )
  })
  
  # Add layout options including the update menus
  p %>%
    layout(
      title = "Volcano Plot",
      xaxis = list(title = "Log2 Fold Change"),
      yaxis = list(title = "-Log10 Adjusted P-value"),
      updatemenus = list(
        list(
          y = 1,  # Position the buttons near the top
          x = 0,   # Center the buttons horizontally
          xanchor = "center",
          yanchor = "top",
          buttons = condition_buttons,
          showactive = TRUE
        )
      )
    )
})
```

### Bar plot
```{r}
renderPlotly({
    req(input$go_1) # Action button
  req(conditions_genes())
  
  # Prepare the data
  data <- conditions_genes()  %>%
        filter(log2fold_change >= input$filter_logfc | log2fold_change <= - input$filter_logfc,
           padj < input$filter_padj) %>% 
    group_by(condition, direction) %>% 
    summarize(n_direct = n()) %>% 
    ungroup()
    
  
  # Create the volcano plot directly with plotly
  plot_ly(data, 
        x = ~condition, 
        y = ~n_direct, 
        color = ~direction, 
        colors = c("#3399FF", "#FF6666"),
        text = ~direction,
        type = 'bar') %>%
  layout(
    title = "DEGs Counts",
    xaxis = list(title = "Condition"),
    yaxis = list(title = "# Genes"),
    barmode = 'group',  
    coloraxis = list(colorbar = list(title = 'Direction'))
  )
})

```


## Row

### Genes table

```{r}

# Reactive expression to load and process the file
conditions_genes <- reactive({
  req(input$conditions_genes)
  
  # Load the file
  conditions_genes <- read.csv(input$conditions_genes$datapath) %>% 
    distinct()
  conditions_genes
})

# Reactive expression to filter significant DEGs
degs_sig <- reactive({
  req(conditions_genes(), input$filter_padj)
  
  degs_sig <- conditions_genes() %>%
    filter(padj < input$filter_padj) %>% 
    distinct()
  
  degs_sig
})

# Reactive expression to further filter DEGs and extract gene names
genes <- reactive({
  req(degs_sig())
  
  genes <- degs_sig() %>%
    filter(log2fold_change >= input$filter_logfc | log2fold_change <= - input$filter_logfc,
           padj < input$filter_padj) %>%
    pull(genes)
  
  genes
})

DT::renderDataTable({
  req(input$go_1) # Action button
  req(degs_sig())
  
  # Define input values for the cutoffs
filter_l2fc <- input$filter_logfc
filter_neg_l2fc <- -input$filter_logfc
  
  # Transform the filtered data into a DataTable
degs_sig() %>%
  datatable(options = list(pageLength = 25)) %>%
  formatStyle(
    'log2fold_change',
    backgroundColor = styleInterval(
      c(filter_neg_l2fc, filter_l2fc),
      c('#3399FF', 'white', '#FF6666')
    ),
    fontWeight = 'bold'  # Optional: make the text bold for better visibility
  )
})

```



